Tutorial: fitting mixed mixture models on text writing data

### 2023-02-07 tutorial/tutorial.Rmd
Update prior

@@

```{r}
data_list <- within( list(), {
  ppts <- ppts  
  nS <- max(ppts) # max no of ppts
  condition <- condition
  K <- max(condition) # no of conditions
  y <- data$iki
  N <- nrow(data)
} )
```

@@

The Stan code can be downloaded from [GitHub](https://osf.io/t34jn/) using `readLines` and `writeLines`. The code will be saved in R's current working directory as "molg.stan" (mixture of log-Gaussians).

```{r eval = F}
# Download Stan code
file <- readLines("https://osf.io/j6dzt/download")
# Save Stan code to current working directory
writeLines(file, "molg.stan")
```

The Stan model can be loaded for data fitting using the `stan_model` function and assign it to the object `mog`. The "MoG.stan" file needs to be available in the current working directory of R or the file path needs to be indicated.

```{r eval = F}
mog <- stan_model(file = "MoG.stan")
```

@@

```{r}
# Parameters to keep in output
pars <- c("beta",                        # fluent typing
          "delta",                       # disfluency slowdown
          "theta",                       # mixing proportion
          "beta_s",                      # by-participant typing speed
          "theta_s",                     # by-participant disfluency probability
          "sigma",                       # variance component
          "sigma_diff", "sigmap_e", "sigma_e",  # variance by mixture component
          "sigma_u",                     # variance for random ppt 
          "log_lik",                     # log likelihood (for model comparison) 
          "y_tilde")                     # predicted data 
```

@@

```{r eval = F}
# Fit model
m <- sampling(mog, 
              data = data_list,
              iter = iterations,
              warmup = warmup,
              chains = n_chains, 
              cores = n_cores,
              pars = pars, # Parameters to keep.
              save_warmup = FALSE, # Don't save the warmup samples.
              seed = 365, 
              control = list(adapt_delta = .96, # default: .9
                             max_treedepth = 12)) # default: 10
```

@@

@@

```{r echo = F}
m <- readRDS("MoG.rda")
```

```{r}
get_elapsed_time(m)/60 # in mins
```

## Reusing model output

The output of the model (i.e. the posterior) can be saved, so we don't need to run the model again (which admittedly can take a while). The function requires the name of the fitted model `m`. We prefer to keep the name of the output similar to the name of the Stan code used. The model is stored as compressed ".rda" file.

```{r eval = F}
saveRDS(m, "MoG.rda", compress = "xz")
```

The model can be load into the environment of R using `readRDS`.

```{r eval = F}
m <- readRDS("MoG.rda")
```

# Model diagnostics

## Convergence

@@

## Posterior predictive check

@@

# Posterior probability distribution

>>>>

The output of the model (i.e. the posterior) can be saved, so we don't need to run the model again (which can take a while). The function requires the name of the fitted model `fit_gaus`. We prefer to keep the name of the output similar to the name of the Stan code used and ideally the R script that is running the model so it is easy to find the script that is running which model. The model is stored as compressed ".rda" file.

@@

```{r}
data_list <- within( list(), {
  ppt <- ppts  
  nS <- max(ppts) # max no of ppts
  condition <- condition
  K <- max(condition) # no of conditions
  y <- data$iki
  N <- nrow(data)
} )
```

@@

The Stan code can be downloaded from [GitHub](https://github.com/jensroes/prowrite-mixture-models/blob/main/tutorial/stan/molg.stan) using `readLines` and `writeLines`. The code will be saved as "molg.stan" (mixture of log-Gaussians) in a directory called "stan" which needs to be available in your current working directory.

```{r eval = F}
# Create an empty directory called "stan"
dir.create("stan")
# Download Stan code
file <- readLines("https://raw.githubusercontent.com/jensroes/prowrite-mixture-models/main/tutorial/stan/molg.stan")
# Save Stan code to current working directory
writeLines(file, "stan/molg.stan")
```

The Stan model can be loaded into R using the `stan_model` function and assign it to the object `molg`. The "molg.stan" file needs to be available in the directory "stan" located inside the current working directory.

```{r eval = F}
molg <- stan_model(file = "stan/molg.stan")
```

@@

```{r}
# Parameters to keep in output
pars <- c("beta",                        # fluent typing
          "delta",                       # disfluency slowdown
          "theta",                       # mixing proportion on logit scale
          "prob",                        # mixing proportion as probability
          "theta_s",                     # by-participant disfluency probability (logit scale)
          "prob_s",                      # by-participant disfluency probability
          "sigma",                       # variance component
          "sigma_diff", "sigmap_e", "sigma_e",  # variance by mixture component
          "sigma_u",                     # variance for random ppt 
          "log_lik",                     # log likelihood (for model comparison) 
          "y_tilde")                     # predicted data 
```

@@

```{r eval = F}
# Fit model
fit_molg <- sampling(molg, 
                     data = data_list,
                     iter = iterations,
                     warmup = warmup,
                     chains = nchains, 
                     cores = ncores,
                     pars = pars, # Parameters to keep.
                     save_warmup = FALSE, # Don't save the warmup samples.
                     seed = 365, 
                     control = list(adapt_delta = 0.96,  # default: 0.9
                                    max_treedepth = 14)) # default: 10
```

@@

```{r eval = F}
saveRDS(object = fit_molg,
        file = "stan/molg.rda", 
        compress = "xz")
```

```{r echo = F}
fit_molg <- readRDS("stan/molg.rda")
```

```{r}
print(fit_molg, pars = c("beta", "delta", "theta"), probs = c(.025, .975))
```

@@

```{r}
get_elapsed_time(fit_molg)/60 # in mins
```

## Model diagnostics

### Convergence

@@

### Posterior predictive check

@@

# Posterior probability distributions
